Gradual Soundness: Lessons from Static Python

### 2022-01-15 pj-7.1/paper.tex
pj: cleanup tables, no landscape

%\overfullrule=1mm

@@

\usepackage{lscape}
\usepackage[framemethod=TikZ]{mdframed}
\newmdenv[roundcorner=2pt,linewidth=0.4pt,linecolor=Black]{spcodebox}
\usepackage{amsthm}
\usepackage{mathpartir}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\usepackage[backend=biber]{biblatex}
\addbibresource{bg.bib}

@@

The tables in this section present fine-grained data for the microbenchmark programs.
Each benchmark has several versions of its code that accomplish the same work:
the original untyped version (\colname{Orig}) and a typed version (\colname{T-Max}) converted for maximum
performance under Static Python and the JIT.
Some benchmarks have a basic typed version (\colname{T-Min}) that adds only
type annotations and necessary casts to the original code,
without using any more advanced Static Python
features such as primitive types or otherwise optimizing the code for Static
Python.
Fannkuch has a second basic version (\colname{T-Min2}) that uses just a few primitive integers
where it can be done without significant changes to the code.

@@

\begin{landscape}

@@

  \begin{tabular}{rrrrrrrrrrrrrrr}
  \colname{T-Max} & \colname{T-Max} & \colname{T-Max} & \colname{T-Max} & \colname{T-Max} & \colname{T-Max} & \colname{T-Min} & \colname{T-Min} & \colname{T-Min} & \colname{T-Min} & \colname{T-Min} & \colname{T-Min} & \colname{Orig} & \colname{Orig} & \colname{Orig} \\
  (\colname{SP} \colname{JIT} \colname{SF}) & (\colname{SP} \colname{JIT}) & (\colname{SP}) & (\colname{JIT} \colname{SF}) & (\colname{JIT}) & () & (\colname{SP} \colname{JIT} \colname{SF}) & (\colname{SP} \colname{JIT}) & (\colname{SP}) & (\colname{JIT} \colname{SF}) & (\colname{JIT}) & () & (\colname{JIT} \colname{SF}) & (\colname{JIT}) & () \\\hline
  $0.57$ & $1.29$ & $7.86$ & $8.36$ & $10.24$ & $16$ & $1.41$ & $2.68$ & $7.67$ & $5.58$ & $7.97$ & $13.32$ & $2.6$ & $3.88$ & $10.17$ \\
  $0.59$ & $1.32$ & $7.92$ & $8.84$ & $9.97$ & $15.73$ & $1.38$ & $2.6$ & $7.9$ & $5.92$ & $7.25$ & $13.59$ & $2.6$ & $4.01$ & $10.18$ \\
  $0.58$ & $1.33$ & $7.89$ & $8.9$ & $10.57$ & $16.2$ & $1.39$ & $2.64$ & $7.74$ & $5.54$ & $7.64$ & $13.83$ & $2.61$ & $4.17$ & $10.73$ \\
  $0.58$ & $1.33$ & $7.92$ & $8.53$ & $10.71$ & $16.71$ & $1.38$ & $2.65$ & $7.8$ & $6.06$ & $7.24$ & $13.95$ & $2.9$ & $3.92$ & $9.96$ \\
  $0.56$ & $1.33$ & $7.83$ & $9.15$ & $9.89$ & $16.62$ & $1.37$ & $2.62$ & $7.68$ & $5.52$ & $7.2$ & $13.5$ & $2.63$ & $3.79$ & $10.04$ \\
  $0.56$ & $1.33$ & $7.97$ & $8.25$ & $10.44$ & $15.65$ & $1.39$ & $2.58$ & $7.8$ & $5.63$ & $7.27$ & $13.52$ & $2.64$ & $3.95$ & $9.99$ \\
  $0.55$ & $1.32$ & $8.18$ & $8.78$ & $10.07$ & $16.22$ & $1.36$ & $2.65$ & $7.79$ & $5.74$ & $7.17$ & $13.34$ & $2.62$ & $3.9$ & $10.16$ \\
  $0.55$ & $1.31$ & $7.85$ & $8.67$ & $10.07$ & $15.73$ & $1.43$ & $2.66$ & $7.82$ & $5.56$ & $7.54$ & $13.68$ & $2.6$ & $3.84$ & $9.69$ \\
  $0.56$ & $1.31$ & $7.96$ & $8.52$ & $10.14$ & $15.82$ & $1.53$ & $2.63$ & $7.9$ & $5.59$ & $6.89$ & $13.79$ & $2.6$ & $4.04$ & $9.82$ \\
  $0.57$ & $1.32$ & $8$ & $8.35$ & $10.57$ & $16.66$ & $1.37$ & $2.71$ & $7.95$ & $5.74$ & $6.91$ & $13.5$ & $2.62$ & $3.85$ & $9.85$ \\
\end{tabular}

\end{table}
\end{landscape}

\begin{landscape}
\begin{table}[tph]
  \caption{Fannkuch microbenchmark data}
  \label{t:mb:fannkuch}
  \footnotesize

  \begin{tabular}{rrrrrrrrrrrrrrr}
  \colname{T-Max} & \colname{T-Max} & \colname{T-Max} & \colname{T-Max} & \colname{T-Max} & \colname{T-Max} & \colname{T-Min} & \colname{T-Min} & \colname{T-Min} & \colname{T-Min} & \colname{T-Min} & \colname{T-Min} & \colname{T-Min2} & \colname{T-Min2} & \colname{T-Min2} \\
  (\colname{SP} \colname{JIT} \colname{SF}) & (\colname{SP} \colname{JIT}) & (\colname{SP}) & (\colname{JIT} \colname{SF}) & (\colname{JIT}) & () & (\colname{SP} \colname{JIT} \colname{SF}) & (\colname{SP} \colname{JIT}) & (\colname{SP}) & (\colname{JIT} \colname{SF}) & (\colname{JIT}) & () & (\colname{SP} \colname{JIT} \colname{SF}) & (\colname{SP} \colname{JIT}) & (\colname{SP}) \\\hline
  $1.27$ & $1.39$ & $34.69$ & $40.74$ & $41$ & $48.23$ & $2.77$ & $2.83$ & $6.18$ & $2.89$ & $2.86$ & $4.07$ & $2.48$ & $2.52$ & $7.28$ \\
  $1.32$ & $1.25$ & $34.35$ & $40.23$ & $40.94$ & $49.03$ & $2.85$ & $2.8$ & $6.12$ & $2.94$ & $2.9$ & $4.06$ & $2.52$ & $2.48$ & $7.28$ \\
  $1.28$ & $1.26$ & $34.67$ & $40.74$ & $40.65$ & $49.42$ & $2.79$ & $2.77$ & $6.38$ & $2.87$ & $2.86$ & $4.1$ & $2.47$ & $2.53$ & $7.24$ \\
  $1.23$ & $1.31$ & $35.71$ & $40.94$ & $39.7$ & $49.35$ & $2.87$ & $2.81$ & $6.23$ & $2.87$ & $2.85$ & $4.29$ & $2.48$ & $2.57$ & $7.35$ \\
  $1.28$ & $1.38$ & $34.85$ & $39.59$ & $40.46$ & $49.18$ & $2.85$ & $2.84$ & $6.17$ & $2.92$ & $2.81$ & $4.14$ & $2.5$ & $2.56$ & $7.34$ \\
  $1.26$ & $1.34$ & $33.96$ & $40.45$ & $39.99$ & $47.79$ & $2.87$ & $2.8$ & $6.21$ & $2.94$ & $2.8$ & $4.06$ & $2.43$ & $2.47$ & $7.17$ \\
  $1.26$ & $1.23$ & $34.65$ & $40.78$ & $40.97$ & $49.7$ & $2.92$ & $2.76$ & $6.04$ & $3.11$ & $3.05$ & $4.08$ & $2.47$ & $2.69$ & $7.18$ \\
  $1.27$ & $1.35$ & $33.83$ & $40.96$ & $41.78$ & $47.92$ & $2.78$ & $2.76$ & $6.13$ & $3.1$ & $2.86$ & $4.19$ & $2.47$ & $2.45$ & $7.14$ \\
  $1.23$ & $1.4$ & $36.23$ & $40.49$ & $40.29$ & $48.68$ & $2.84$ & $2.76$ & $6.13$ & $2.87$ & $2.96$ & $4.31$ & $2.43$ & $2.55$ & $7.23$ \\
  $1.23$ & $1.27$ & $36.05$ & $40.72$ & $40.11$ & $48.29$ & $2.92$ & $2.79$ & $6.17$ & $2.91$ & $2.96$ & $4.1$ & $2.52$ & $2.49$ & $7.2$ \\
\end{tabular}

\begin{tabular}{rrrrrr}
  \colname{T-Min2} & \colname{T-Min2} & \colname{T-Min2} & \colname{Orig} & \colname{Orig} & \colname{Orig} \\
  (\colname{JIT} \colname{SF}) & (\colname{JIT}) & () & (\colname{JIT} \colname{SF}) & (\colname{JIT}) & () \\\hline
  $3.16$ & $3.08$ & $4.59$ & $2.81$ & $2.94$ & $4.02$ \\
  $3.26$ & $3.08$ & $4.51$ & $2.66$ & $2.85$ & $4.22$ \\
  $3.11$ & $3.18$ & $4.59$ & $2.69$ & $2.78$ & $4.01$ \\
  $3.18$ & $3.14$ & $4.73$ & $2.75$ & $2.8$ & $4.01$ \\
  $3.14$ & $3.12$ & $4.63$ & $2.75$ & $2.92$ & $4.07$ \\
  $3.23$ & $3.09$ & $4.53$ & $2.78$ & $2.83$ & $4.36$ \\
  $3.41$ & $3.21$ & $4.48$ & $2.84$ & $2.81$ & $4.19$ \\
  $3.43$ & $3.14$ & $4.53$ & $2.7$ & $2.86$ & $4.25$ \\
  $3.08$ & $3.24$ & $4.53$ & $2.76$ & $2.78$ & $4.08$ \\
  $3.16$ & $3.18$ & $4.45$ & $2.89$ & $2.76$ & $4.16$ \\
\end{tabular}
\end{table}

\end{landscape}

@@

  Fannkuch \colname{T-Min2} & \texttt{time ./python -X jit -X jit-list-file=Tools/benchmarks/jitlist\_fannkuch\_static\_basic2.txt -X jit-enable-jit-list-wi} \\

@@

  Fannkuch \colname{T-Min2} & \texttt{time ./python -X install-strict-loader Tools/benchmarks/fannkuch\_static\_basic2.py 5} \\

@@

  Fannkuch \colname{T-Min2} & \texttt{time ./python Tools/benchmarks/fannkuch\_static\_basic2.py 5} \\

>>>>

\overfullrule=1mm

@@

\usepackage[framemethod=TikZ]{mdframed}
\newmdenv[roundcorner=2pt,linewidth=0.4pt,linecolor=Black]{spcodebox}
\usepackage{amsthm}
\usepackage{mathpartir}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\usepackage[backend=biber]{biblatex}
\addbibresource{bg.bib}

@@

The tables in this section present fine-grained data for the microbenchmark programs.
Each benchmark has several versions of its code that accomplish the same work:
the original untyped version (\colname{Orig}) and a typed version (\colname{T-Max}) converted for maximum
performance under Static Python and the JIT.
Some benchmarks have a basic typed version (\colname{T-Min}) that adds only
type annotations and necessary casts to the original code,
without using any more advanced Static Python
features such as primitive types or otherwise optimizing the code for Static
Python.
Fannkuch has a second basic version (\colname{T-Min-2}) that uses just a few primitive integers
where it can be done without significant changes to the code.

@@

@@

\begin{tabular}{rrrrrr}
  \colname{T-Max} & \colname{T-Max} & \colname{T-Max} & \colname{T-Max} & \colname{T-Max} & \colname{T-Max} \\
  (\colname{SP} \colname{JIT} \colname{SF}) & (\colname{SP} \colname{JIT}) & (\colname{SP}) & (\colname{JIT} \colname{SF}) & (\colname{JIT}) & () \\\hline
  $0.57$ & $1.29$ & $7.86$ & $8.36$ & $10.24$ & $16$ \\
  $0.59$ & $1.32$ & $7.92$ & $8.84$ & $9.97$ & $15.73$ \\
  $0.58$ & $1.33$ & $7.89$ & $8.9$ & $10.57$ & $16.2$ \\
  $0.58$ & $1.33$ & $7.92$ & $8.53$ & $10.71$ & $16.71$ \\
  $0.56$ & $1.33$ & $7.83$ & $9.15$ & $9.89$ & $16.62$ \\
  $0.56$ & $1.33$ & $7.97$ & $8.25$ & $10.44$ & $15.65$ \\
  $0.55$ & $1.32$ & $8.18$ & $8.78$ & $10.07$ & $16.22$ \\
  $0.55$ & $1.31$ & $7.85$ & $8.67$ & $10.07$ & $15.73$ \\
  $0.56$ & $1.31$ & $7.96$ & $8.52$ & $10.14$ & $15.82$ \\
  $0.57$ & $1.32$ & $8$ & $8.35$ & $10.57$ & $16.66$ \\
\end{tabular}

\begin{tabular}{rrrrrr}
  \colname{T-Min} & \colname{T-Min} & \colname{T-Min} & \colname{T-Min} & \colname{T-Min} & \colname{T-Min} \\
  (\colname{SP} \colname{JIT} \colname{SF}) & (\colname{SP} \colname{JIT}) & (\colname{SP}) & (\colname{JIT} \colname{SF}) & (\colname{JIT}) & () \\\hline
  $1.41$ & $2.68$ & $7.67$ & $5.58$ & $7.97$ & $13.32$ \\
  $1.38$ & $2.6$ & $7.9$ & $5.92$ & $7.25$ & $13.59$ \\
  $1.39$ & $2.64$ & $7.74$ & $5.54$ & $7.64$ & $13.83$ \\
  $1.38$ & $2.65$ & $7.8$ & $6.06$ & $7.24$ & $13.95$ \\
  $1.37$ & $2.62$ & $7.68$ & $5.52$ & $7.2$ & $13.5$ \\
  $1.39$ & $2.58$ & $7.8$ & $5.63$ & $7.27$ & $13.52$ \\
  $1.36$ & $2.65$ & $7.79$ & $5.74$ & $7.17$ & $13.34$ \\
  $1.43$ & $2.66$ & $7.82$ & $5.56$ & $7.54$ & $13.68$ \\
  $1.53$ & $2.63$ & $7.9$ & $5.59$ & $6.89$ & $13.79$ \\
  $1.37$ & $2.71$ & $7.95$ & $5.74$ & $6.91$ & $13.5$ \\
\end{tabular}

\begin{tabular}{rrr}
  \colname{Orig} & \colname{Orig} & \colname{Orig} \\
  (\colname{JIT} \colname{SF}) & (\colname{JIT}) & () \\\hline
  $2.6$ & $3.88$ & $10.17$ \\
  $2.6$ & $4.01$ & $10.18$ \\
  $2.61$ & $4.17$ & $10.73$ \\
  $2.9$ & $3.92$ & $9.96$ \\
  $2.63$ & $3.79$ & $10.04$ \\
  $2.64$ & $3.95$ & $9.99$ \\
  $2.62$ & $3.9$ & $10.16$ \\
  $2.6$ & $3.84$ & $9.69$ \\
  $2.6$ & $4.04$ & $9.82$ \\
  $2.62$ & $3.85$ & $9.85$ \\
\end{tabular}

\begin{table}[tph]
  \caption{Fannkuch microbenchmark data}
  \label{t:mb:fannkuch}
  \footnotesize

  \begin{tabular}{rrrrrr}
  \colname{T-Max} & \colname{T-Max} & \colname{T-Max} & \colname{T-Max} & \colname{T-Max} & \colname{T-Max} \\
  (\colname{SP} \colname{JIT} \colname{SF}) & (\colname{SP} \colname{JIT}) & (\colname{SP}) & (\colname{JIT} \colname{SF}) & (\colname{JIT}) & () \\\hline
  $1.27$ & $1.39$ & $34.69$ & $40.74$ & $41$ & $48.23$ \\
  $1.32$ & $1.25$ & $34.35$ & $40.23$ & $40.94$ & $49.03$ \\
  $1.28$ & $1.26$ & $34.67$ & $40.74$ & $40.65$ & $49.42$ \\
  $1.23$ & $1.31$ & $35.71$ & $40.94$ & $39.7$ & $49.35$ \\
  $1.28$ & $1.38$ & $34.85$ & $39.59$ & $40.46$ & $49.18$ \\
  $1.26$ & $1.34$ & $33.96$ & $40.45$ & $39.99$ & $47.79$ \\
  $1.26$ & $1.23$ & $34.65$ & $40.78$ & $40.97$ & $49.7$ \\
  $1.27$ & $1.35$ & $33.83$ & $40.96$ & $41.78$ & $47.92$ \\
  $1.23$ & $1.4$ & $36.23$ & $40.49$ & $40.29$ & $48.68$ \\
  $1.23$ & $1.27$ & $36.05$ & $40.72$ & $40.11$ & $48.29$ \\
\end{tabular}

\begin{tabular}{rrrrrr}
  \colname{T-Min} & \colname{T-Min} & \colname{T-Min} & \colname{T-Min} & \colname{T-Min} & \colname{T-Min} \\
  (\colname{SP} \colname{JIT} \colname{SF}) & (\colname{SP} \colname{JIT}) & (\colname{SP}) & (\colname{JIT} \colname{SF}) & (\colname{JIT}) & () \\\hline
  $2.77$ & $2.83$ & $6.18$ & $2.89$ & $2.86$ & $4.07$ \\
  $2.85$ & $2.8$ & $6.12$ & $2.94$ & $2.9$ & $4.06$ \\
  $2.79$ & $2.77$ & $6.38$ & $2.87$ & $2.86$ & $4.1$ \\
  $2.87$ & $2.81$ & $6.23$ & $2.87$ & $2.85$ & $4.29$ \\
  $2.85$ & $2.84$ & $6.17$ & $2.92$ & $2.81$ & $4.14$ \\
  $2.87$ & $2.8$ & $6.21$ & $2.94$ & $2.8$ & $4.06$ \\
  $2.92$ & $2.76$ & $6.04$ & $3.11$ & $3.05$ & $4.08$ \\
  $2.78$ & $2.76$ & $6.13$ & $3.1$ & $2.86$ & $4.19$ \\
  $2.84$ & $2.76$ & $6.13$ & $2.87$ & $2.96$ & $4.31$ \\
  $2.92$ & $2.79$ & $6.17$ & $2.91$ & $2.96$ & $4.1$ \\
\end{tabular}

\begin{tabular}{rrrrrr}
  \colname{T-Min-2} & \colname{T-Min-2} & \colname{T-Min-2} & \colname{T-Min-2} & \colname{T-Min-2} & \colname{T-Min-2} \\
  (\colname{SP} \colname{JIT} \colname{SF}) & (\colname{SP} \colname{JIT}) & (\colname{SP}) & (\colname{JIT} \colname{SF}) & (\colname{JIT}) & () \\\hline
  $2.48$ & $2.52$ & $7.28$ & $3.16$ & $3.08$ & $4.59$ \\
  $2.52$ & $2.48$ & $7.28$ & $3.26$ & $3.08$ & $4.51$ \\
  $2.47$ & $2.53$ & $7.24$ & $3.11$ & $3.18$ & $4.59$ \\
  $2.48$ & $2.57$ & $7.35$ & $3.18$ & $3.14$ & $4.73$ \\
  $2.5$ & $2.56$ & $7.34$ & $3.14$ & $3.12$ & $4.63$ \\
  $2.43$ & $2.47$ & $7.17$ & $3.23$ & $3.09$ & $4.53$ \\
  $2.47$ & $2.69$ & $7.18$ & $3.41$ & $3.21$ & $4.48$ \\
  $2.47$ & $2.45$ & $7.14$ & $3.43$ & $3.14$ & $4.53$ \\
  $2.43$ & $2.55$ & $7.23$ & $3.08$ & $3.24$ & $4.53$ \\
  $2.52$ & $2.49$ & $7.2$ & $3.16$ & $3.18$ & $4.45$ \\
\end{tabular}

\begin{tabular}{rrr}
  \colname{Orig} & \colname{Orig} & \colname{Orig} \\
  (\colname{JIT} \colname{SF}) & (\colname{JIT}) & () \\\hline
  $2.81$ & $2.94$ & $4.02$ \\
  $2.66$ & $2.85$ & $4.22$ \\
  $2.69$ & $2.78$ & $4.01$ \\
  $2.75$ & $2.8$ & $4.01$ \\
  $2.75$ & $2.92$ & $4.07$ \\
  $2.78$ & $2.83$ & $4.36$ \\
  $2.84$ & $2.81$ & $4.19$ \\
  $2.7$ & $2.86$ & $4.25$ \\
  $2.76$ & $2.78$ & $4.08$ \\
  $2.89$ & $2.76$ & $4.16$ \\
\end{tabular}

@@

  Fannkuch \colname{T-Min-2} & \texttt{time ./python -X jit -X jit-list-file=Tools/benchmarks/jitlist\_fannkuch\_static\_basic2.txt -X jit-enable-jit-list-wi} \\

@@

  Fannkuch \colname{T-Min-2} & \texttt{time ./python -X install-strict-loader Tools/benchmarks/fannkuch\_static\_basic2.py 5} \\

@@

  Fannkuch \colname{T-Min-2} & \texttt{time ./python Tools/benchmarks/fannkuch\_static\_basic2.py 5} \\
